\section{Introduction}
The Fr\'echet distance was introduced by Maurice Fr\'echet in 1906 \citep{frechet06} and is well known as a fundamental metric in abstract spaces in the field of mathematics and computational geometry.
Over the years, it has found applications in various domains, e.g., as a distance measure between probability distributions it is equivalent to the Wasserstein-2 distance \citep{kantorovich60,wasserstein69} for the $\ell^2$-norm.
\citet{dowson82} demonstrated that for normal distributions the Fr\'echet distance can be calculated explicitly and since then it is used a basis of various evaluation metrics such as the Fr\'echet inception distance \citep{heusel17}, Fr\'echet audio distance \citep{kilgour19}, or the Fr\'echet ChemNet distance \citep{preuer18}.
In the early 1990s, Alt and Godau were the first to apply the Fr\'echet distance in measuring the similarity of polygonal curves \citep{alt92,alt95}.
However, there is no explicit way to find the Fr\'echet distance in this setting directly.
Instead, the problem can be solved with dynamic programming and finding efficient algorithms remains a vibrant area of research \citep{aronov06,avraham14,agarwal14,bringmann19}.
Our contribution to this ongoing exploration is an efficient algorithm for the discrete Fr\'echet distance that is optimized to run fast on modern hardware.
In particular, we reformulate the recursive algorithm proposed by \citet{eiter94} as an iterative, branchless algorithm and reduce its quadratic memory requirement to a linear one.
Our formulation requires just four lines of pseudo-code, excluding function signatures, and can be vectorized to take full advantage of modern CPU and GPU architectures.
To the best of our knowledge, we are the first to comprehensively propose these modifications although it is possible that others---while only referring to \citet{eiter94}---are already using some of the modifications in practice without explicitly noting.

The remainder of this paper is structured as follows:
First, we describe the continuous and discrete Fr\'echet distance for polygonal curves in Sec.~\ref{sec:frechet}.
In Sec.~\ref{sec:fast} we present our improved algorithm.
Subsequently, we discuss its relation to dynamic time wrapping (DTW) and the Levenshtein distance in Sec.~\ref{sec:dtw_levenshtein}.
Finally, we outline how our solution can be parallelized in Sec.~\ref{sec:parallel}.

\section{The Fr\'echet Distance}
\label{sec:frechet}
Formally, let $p$ and $q$ be two curves in a metric space $\mathcal{S}$.
Then, the (continuous) Fr\'echet distance $\delta_\mathrm{F}$ between $p$ and $q$ is defined as the infimum over all possible continuous, non-decreasing, surjections $\alpha: [0, 1] \to [0, 1]$ and $\beta: [0, 1] \to [0, 1]$ of the maximum over all $t \in [0, 1]$ of the distance in $\mathcal{S}$ between $p(\alpha(t))$ and $q(\beta(t))$,
\begin{equation*}
    \delta_\mathrm{F}(p, q) = \inf_{\alpha,\beta} \; \max_{t \in [0, 1]} \left\| p(\alpha(t)) - q(\beta(t)) \right\|_\mathcal{S},
\end{equation*}
where $\| \bullet \|_\mathcal{S}$ is the distance function of $\mathcal{S}$.
Informally, this measure is often described as the minimal leash length, measured by $\| \bullet \|_\mathcal{S}$, of a man walking his dog if both are moving on trajectories $p(\alpha(t))$ and $q(\beta(t))$, respectively, where $t$ is interpreted as a measure of time.
In this interpretation the restriction that $\alpha(t)$ and $\beta(t)$ be non-decreasing means that neither the dog nor its owner can backtrack.
Adding this constraint makes the Fr\'echet distance unconditionally stronger than the Hausdorff distance in the following sense: the Fr\'echet distance between two curves is always greater or equal than their corresponding Hausdorff distance.

Although non-straightforward to compute, the Fr\'echet distance has been used successfully in various fields such as curve simplification \citep{agarwal05,bereg08,kreveld18,kerkhof19}, map-matching \citep{alt03a,driemel12,sharma17} and clustering \citep{buchin08b,besse16,buchin19b,buchin19c}.
The Fr\'echet distance also has applications in matching biological sequences \citep{wylie12}, analysing tracking data \citep{brakatsoulas05,buchin08a}, and matching coastline data \citep{mascret06}.
In particular for clustering, a major advantage of the Fr\'echet distance is that it fulfills the triangle inequality and therefore is not only a distance measure but also a metric. 
Other measures of curve similarity, such as DTW \citep{vintsyuk68,sankoff83}, do not obey the triangle inequality which limits their application in practice.

Computing the Fr\'echet distance between two precise curves can be done in near-quadratic time \citep{agarwal14,alt95,buchin17}, and assuming the strong exponential time hypothesis, it cannot be computed or even approximated well in strongly subquadratic time unless SETH fails \citep{bringmann14,buchin19a}.
In practice, algorithms with a theoretical quadratic complexity often rely on certain pre-processing steps and are complex to implement in general \citep{bringmann19}.
Furthermore, to the best of our knowledge, almost all of the analyses and applications of the continuous Fr\'echet distance either explicitly or implicitly rely on the Euclidean distance for $\| \bullet \|_\mathcal{S}$ which makes the corresponding pathing in \textit{free-space} a convex optimization problem \citep{alt92,alt95}.
Extending the results to more general metrics is not straightforward as one quickly looses auxiliary properties such as the convexity in \textit{free-space}.

One way to overcome these challenges is to estimate the continuous Fr\'echet distance of polygonal curves $\delta_\mathrm{F}$ with a discrete approximation $\delta_\mathrm{dF}$:
Instead of taking all points of the (continuous) curves into account, the discrete Fr\'echet distance only includes a finite number of them.
A typical choice for the set of the points is the set of vertices of the given polygonal curve itself, however, in order to decrease the approximation error, additional points can be sampled between them.
Informally and in comparison to the common \textit{man-walking-his-dog} analogy, the discrete Fr\'echet distance can be thought of as the minimal leash length between two frogs jumping between stones, where \textit{stones} are the vertices and the leash is not taken into consideration during jumps.

In 1994, \citet{eiter94} defined the discrete Fr\'echet distance $\delta_\mathrm{dF}$ between two $D$-dimensional polygonal chains $p \in \mathbb{R}^{P \times D}$ and $q \in \mathbb{R}^{Q \times D}$, proposed the simple Alg.~\ref{alg:vanilla} of complexity $\mathcal{O}(D'PQ)$, where $D'$ is the complexity of evaluating $f$, and showed that the difference between the continuous Fr\'echet distance and its discrete variant is bound by the sample width of the curves,
\begin{equation*}
    \delta_\mathrm{F}(p, q) \;\le\; \delta_\mathrm{dF}(p, q) \;\le\; \delta_\mathrm{F}(p, q) + \max\{ \varepsilon_p, \varepsilon_q \},
\end{equation*}
where $\varepsilon_p$ and $\varepsilon_q$ are the largest distance between adjacent points (the \textit{stones}) on $p$ and $q$, respectively.

\begin{algorithm}
    \caption{The original algorithm for calculating the discrete Fr\'echet distance as proposed by \citet{eiter94} for a generic distance measure $f: \mathbb{R}^D \times \mathbb{R}^D \mapsto \mathbb{R}$.}
    \label{alg:vanilla}
    \begin{algorithmic}[1]
        \Function{fr\'echet\_distance}{$p$: $\mathbb{R}^{P \times D}$, $q$: $\mathbb{R}^{Q \times D}$, $f$: $\mathbb{R}^D \times \mathbb{R}^D \mapsto \mathbb{R}$} $\to \mathbb{R}$
            \State $M: \mathbb{R}^{P \times Q}$ \label{alg:vanilla:M}
            \Statex
            \Function{eval}{$i: \mathbb{N}$, $j: \mathbb{N}$} $\to \mathbb{R}$
                \If{$M_{ij} > -1$}
                    \State \Return $M_{ij}$
                \EndIf
                \Statex
                \State $d: \mathbb{R}$ $\gets$ \Call{$f$}{$p_i, q_j$}
                \If{$i = 1 \;\land\; j = 1$} \label{alg:vanilla:if1}
                    \State $M_{ij} \gets d$
                \ElsIf{$i > 1 \;\land\; j = 1$} \label{alg:vanilla:if2}
                    \State $M_{ij} \gets \max\{\Call{eval}{i - 1, 1},\, d\}$
                \ElsIf{$i = 1 \;\land\; j > 1$} \label{alg:vanilla:if3}
                    \State $M_{ij} \gets \max\{\Call{eval}{1, j - 1},\, d\}$
                \Else \label{alg:vanilla:if4}
                    \State $M_{ij} \gets \max\{\min\{\Call{eval}{i - 1, j},\, \Call{eval}{i - 1, j - 1},\, \Call{eval}{i, j - 1}\},\, d\}$
                \EndIf
                \Statex
                \State \Return $M_{ij}$
            \EndFunction
            \Statex
            \State \Return \Call{eval}{$P, Q$}
        \EndFunction
    \end{algorithmic}
\end{algorithm}

Since then, the discrete Fr\'echet distance is used frequently, e.g., \citet{sriraghavendra07} have used it for handwriting recognition and \citet{jiang08} used the discrete Fr\'echet distance to tackle the protein structure-structure alignment problem.
Arguably, for the latter utilizing the discrete Fr\'echet distance even makes more sense than the continuous Fr\'echet distance as the backbone of a protein is simply a polygonal chain in 3D, with each vertex being the alpha-carbon atom of a residue.
Hence, if the continuous Fr\'echet distance is realized by an alpha-carbon atom and some other point which does not represent an atom, it is not meaningful biologically.
Later, \citet{zhu07} extended this work and analyzed the protein local structural alignment problem using bounded discrete Fr\'echet distance.

Furthermore, the Fr\'echet distance serves as a pivotal metric for assessing similarities among vessel trajectories in maritime research.
Here, the discrete Fr\'echet distance again emerges as a more fitting measure in this context compared to its continuous variant, aligning seamlessly with the nature of trajectory data derived from discrete GNSS updates.
If not used for benchmarking alternative approaches \citep{li20,liu23,luo23}, the discrete Fr\'echet distance is used a fundamental building block for movement analytics, clustering, and classification of the trajectories, for example:
\citet{sakan23} studied route characteristics by using the discrete Fr\'{e}chet distance to measure the similarity of individual container fleets from a statistical perspective.
\citet{cao18} used the discrete Fr\'{e}chet distance to solve the maximum distance between vessel trajectories and obtained a distance matrix, which was then decomposed using PCA to determine trajectory cluster.
Similarly, \citet{roberts18} constructed a graph where vessel trajectories are the vertices and the edges are weighted by the discrete Fr\'echet distance between two vertices.
Subsequently, a minimum spanning tree was created from this graph and used for clustering.

Although applied successfully in practice, the algorithm by \citet{eiter94} has three major disadvantages:
First, the recursive formulation makes the algorithm impractical for larger curves as the recursive calls cannot be elided by tail recursions, thus making the algorithm slow and prone to stack overflows.
Secondly, the algorithm needs to allocate a memory block $M$ of size $P \times Q$ (see line~\ref{alg:vanilla:M} in Alg.~\ref{alg:vanilla}).
For large curves this allocation is slow, pollutes cache lines, and potentially constrains parallelization due to limited memory resources.
Thirdly, the explicit branching points on lines~\ref{alg:vanilla:if1}, \ref{alg:vanilla:if2}, \ref{alg:vanilla:if3} and \ref{alg:vanilla:if4} in Alg.~\ref{alg:vanilla} put stress on the branch predictor and make it hard to reformulate the algorithm for an effective use on \textit{single instruction, multiple data} (SIMD) or \textit{single instruction, multiple threads} (SIMT) architectures without causing significant branch divergences.

\section{An Iterative, Linear-Memory Algorithm}
\label{sec:fast}
In this section we introduce a new algorithm for finding the discrete Fr\'echet distance between two polygonal curves $p \in \mathbb{R}^{P \times D}$ and $q \in \mathbb{R}^{Q \times D}$.
Our algorithm uses (left) fold/reduce operators for the binary functions \textsc{fr\'echet\_maxmin} and $\max$ on lines \ref{alg:fast:scan1} and \ref{alg:fast:scan2}, and a (left) scan/accumulate operator for the binary function \textsc{fr\'echet\_next} on line~\ref{alg:fast:fold}, where we borrowed the notation for the operators, $/$ and $\backslash$, from APL \citep{iverson62,iverson79}.
In particular, note that our scan operator prepends the initial value $\max\{ a_1, x_1 \}$ to the returned sequence.
On line \ref{alg:fast:scan2} the parameters of the binary function $\max$ have the same type and thus we implicitly take the first value of the passed sequence as the initial value and pass the remaining sequence as the second argument.
Pseudocodes for the operators are shown in the appendix in Alg.~\ref{alg:scan} and Alg.~\ref{alg:fold}.

\begin{theorem}[Fast Discrete Fr\'echet Algorithm]
\label{thm:fast}
Alg.~\ref{alg:fast} calculates the discrete Fr\'echet distance given a distance matrix $d \in \mathbb{R}^{P \times Q}$.
The algorithms iteratively consumes the rows of $d$ such that each row, or even each element, can be computed lazily; the memory requirement is therefore reduced to $\mathcal{O}(2Q)$.
The complexity of Alg.~\ref{alg:fast} is $\mathcal{O}(PQ)$ if $d$ is precomputed and $\mathcal{O}(D'PQ)$ if $d$ is evaluated lazily, where $D'$ is the complexity of evaluating a single element of $d$.

\begin{algorithm}
    \caption{A fast and concise algorithm that uses the fold of \textsc{fr\'echet\_next} (line~\ref{alg:fast:fold}) and the scans of \textsc{fr\'echet\_maxmin} (line~\ref{alg:fast:scan1}) and $\max$ (line~\ref{alg:fast:scan2}) to map a given distance matrix $d_{ij} = \| p_i - q_j \|_\mathcal{S}$ to the discrete Fr\'echet distance of $p \in \mathbb{R}^{P \times D}$ and $q \in \mathbb{R}^{Q \times D}$. The rows of $d$ can be evaluated lazily to achieve the linear memory requirement. The minimum on line~\ref{alg:fast:min} is taken element-wise. On line~\ref{alg:fast:scan1} the column vectors $a_{2 \ldots Q}$ and $x_{2 \ldots Q}$ are stacked into a $\mathbb{R}^{(Q - 1) \times 2}$ matrix and iterated row-wise.}
    \label{alg:fast}
    \begin{algorithmic}[1]
        \Function{fr\'echet\_maxmin}{$a: \mathbb{R}$, $x: \mathbb{R}^2$} $\to \mathbb{R}$
            \State \Return $\max\{ \min\{ a,\, x_1 \},\, x_2 \}$
        \EndFunction
        \Statex
        \Function{fr\'echet\_next}{$a: \mathbb{R}^Q$, $x: \mathbb{R}^{Q}$} $\to \mathbb{R}^Q$
            \State $a_{2 \ldots Q} \gets \min\{ a_{1 \ldots Q-1},\, a_{2 \ldots Q} \}$ \label{alg:fast:min}
            \State \Return \Call{fr\'echet\_maxmin$\backslash$}{$\max\{a_1, x_1\},\, [a_{2 \ldots Q} \;|\; x_{2 \ldots Q}]$} \label{alg:fast:scan1}
        \EndFunction
        \Statex
        \Function{fr\'echet\_distance}{$d: \mathbb{R}^{P \times Q}$} $\to \mathbb{R}$
            \State \Return $[$\Call{fr\'echet\_next$/$}{$\max\!\backslash(d_1), d_{2 \ldots P}$}$]_Q$ \label{alg:fast:fold} \label{alg:fast:scan2}
        \EndFunction
    \end{algorithmic}
\end{algorithm}
\end{theorem}

\begin{proof}[Proof of Theorem~\ref{thm:fast}]
In order to prove Theorem~\ref{thm:fast} we start by rewriting Alg.~\ref{alg:vanilla} and replace the recursive calls with two explicit iterations over the points in $p$ and $q$.
Alg.~\ref{alg:vanilla} finds the discrete Fr\'echet distance between $p$ and $q$ \textit{top-down} by recursively solving the dynamic programming problem
\begin{equation}
    \label{eq:dynprogprob}
    M_{ij} = \max\{ \min\{ M_{i-1,j}, M_{i-1,j-1}, M_{i,j-1} \}, d_{ij} \},
\end{equation}
where $\delta_\mathrm{dF} = M_{PQ}$.
Alg.~\ref{alg:no_recursion} shows an algorithm that also solves Eq.~\eqref{eq:dynprogprob} but with a \textit{bottom-up} approach.
The nested loops start at low indices and eventually accumulate the result in $M_{PQ}$ until $i=P$ and $j=Q$.
In fact, at every point $M_{\alpha \beta}$ is the discrete Fr\'echet distance for the polygonal curves $p_i$ and $q_j$ with $i \in [1, \alpha]$ and $j \in [1, \beta]$, respectively.
The complexity of this algorithm is $\mathcal{O}(D'PQ)$ where the complexity of evaluating $f(p_i, q_j)$ on line \ref{alg:no_recursion:d} is $\mathcal{O}(D')$.

We believe that at least this simplification is already well established since \citet{ahn10} used a decision algorithm that is a close adaptation of Alg.~\ref{alg:vanilla} and is already implemented without recursions---yet still with a quadratic memory requirement and branches.
In Sec.~\ref{sec:dtw_levenshtein} we will discuss the relation of the discrete Fr\'echet distance to DTW and the Levenshtein distance.
Here as well iterative algorithms without recursions are common.

\begin{algorithm}
    \caption{Variant of Alg.~\ref{alg:vanilla} without recursions.}
    \label{alg:no_recursion}
    \begin{algorithmic}[1]
        \Function{fr\'echet\_distance}{$p: \mathbb{R}^{P \times D}$, $q: \mathbb{R}^{Q \times D}$, $f: \mathbb{R}^D \times \mathbb{R}^D \mapsto \mathbb{R}$} $\to \mathbb{R}$
            \State $M: \mathbb{R}^{P \times Q}$
            \Statex
            \For{$i \gets 1,P$}
                \For{$j \gets 1,Q$}
                    \State $d: \mathbb{R} \gets f(p_i, q_j)$ \label{alg:no_recursion:d}
                    \Statex
                    \If{$i = 1 \;\land\; j = 1$}
                        \State $M_{11} \gets d$
                    \ElsIf{$i > 1 \;\land\; j = 1$}
                        \State $M_{i,1} \gets \max\{M_{i-1,1},\, d \}$
                    \ElsIf{$i = 1 \;\land\; j > 1$}
                        \State $M_{1,j} \gets \max\{M_{1,j-1},\, d \}$
                    \Else
                        \State $M_{ij} \gets \max\{\min\{M_{i-1,j},\, M_{i-1,j-1},\, M_{i,j-1}\},\, d \}$
                    \EndIf
                \EndFor
            \EndFor
            \Statex
            \State \Return $M_{PQ}$   
        \EndFunction
    \end{algorithmic}
\end{algorithm}

Next, we remove branching points by evaluating the elements of $M_{ij}$ where either $i=1$ or $j=1$ before entering the nested loops with scans/accumulates of the first column (line~\ref{alg:inplace:scan1} of Alg.~\ref{alg:inplace}) and first row (line~\ref{alg:inplace:scan2} of Alg.~\ref{alg:inplace}) of $d$ using the $\max$ function, respectively.
In fact, this variant can be formulated as an in-place algorithm that operates on the distance matrix $d \in \mathbb{R}^{P \times Q}$ of $p$ and $q$ without needing any further allocations.
We show this variant in Alg.~\ref{alg:inplace}.

\begin{algorithm}
    \caption{In-place variant of Alg.~\ref{alg:vanilla}. This variant directly maps a given distance matrix $d_{ij} = \|p_i - q_j\|_\mathcal{S}$  to the corresponding Fr\'echet distance. Note that this variant has no explicit branching points.}
    \label{alg:inplace}
    \begin{algorithmic}[1]
        \Function{fr\'echet\_distance}{$d: \mathbb{R}^{P \times Q}$} $\to \mathbb{R}$
            \State $d_{:,1} \gets \max\!\backslash(d_{:,1})$ \label{alg:inplace:scan1}
            \State $d_{1,:} \gets \max\!\backslash(d_{1,:})$ \label{alg:inplace:scan2}
            \Statex
            \For{$i \gets 2,P$}
                \For{$j \gets 2,Q$}
                    \State $d_{ij} \gets \max\{\min\{d_{i-1,j},\, d_{i-1,j-1},\, d_{i,j-1}\},\, d_{ij} \}$
                \EndFor
            \EndFor
            \Statex
            \State \Return $d_{PQ}$   
        \EndFunction
    \end{algorithmic}
\end{algorithm}

Finally, we note that only two adjacent rows of $M$ (or $d$ in Alg.~\ref{alg:inplace}) are needed during the iterations.
In Alg.~\ref{alg:linear} we extract these rows into $u,v \in \mathbb{R}^Q$ and thus reduce the overall memory requirement to $\mathcal{O}(2Q)$.
In Alg.~\ref{alg:fast} the iterations over $p_i$ and $q_j$ have been rewritten as a fold/reduce and a scan/accumulate operation, respectively.

\begin{algorithm}
    \caption{Variant of Alg.~\ref{alg:vanilla} without recursions and explicit branching points, and a linear memory requirement.}
    \label{alg:linear}
    \begin{algorithmic}[1]
        \Function{fr\'echet\_distance}{$p: \mathbb{R}^{P \times D}$, $q: \mathbb{R}^{Q \times D}$, $f: \mathbb{R}^D \times \mathbb{R}^D \mapsto \mathbb{R}$} $\to \mathbb{R}$
            \State $v: \mathbb{R}^Q \gets \max\!\backslash(f(p_1, q))$
            \Statex
            \For{$i \gets 2,P$}
                \State $u: \mathbb{R}^{Q-1} \gets \min\{ v_{1 \ldots Q-1},\, v_{2 \ldots Q} \}$
                \Statex
                \State $v_1 \gets \max\{ v_1, f(p_i, q_1) \}$
                \For{$j \gets 2,Q$}
                    \State $v_j \gets \max\{ \min\{ u_{j-1},\, v_{j-1} \},\, f(p_i, q_j) \}$ 
                \EndFor
            \EndFor
            \Statex
            \State \Return $v_Q$
        \EndFunction
    \end{algorithmic}
\end{algorithm}
\end{proof}

The quadratic complexity of the algorithms can be improved if further approximation are made.
For example, \citet{aronov06} presented an efficient approximation algorithms for computing the discrete Fr\'echet distance of two natural classes of curves: $\kappa$-bounded curves and backbone curves.
They also proposed a pseudo-output-sensitive algorithm for computing the discrete Fr\'echet distance exactly.
However, even though the discrete Fr\'echet distance can be calculated faster for several restricted versions \citep{driemel12,gudmundsson18,buchin10,maheshwari11} our proposed algorithm works for the general case.
In particular, it does not make any assumptions about the metric or trajectory properties, e.g., Sakoe-Chiba bands, and scaling to higher dimensions only depends on $D'$.

\section{DTW and the Levenshtein Distance}
\label{sec:dtw_levenshtein}
Fr\'echet: \textit{metric} result / type of $d$ has to be a set with a \textit{total preorder} that induces min/max. Fr\'echet algorithm is a search algorithm.

In contrast to other measures of curve similarity, such as dynamic time warping (DTW) \citep{vintsyuk68,sankoff83} or the average Fr\'echet distance \citep{brakatsoulas05}, the discrete and continuous Fr\'echet distance fulfill the triangle inequality 

Levenshtein distance: introduced by Vladimir Levenshtein \citep{levenshtein65}, matrix implementation \citep{wagner74}, fast: \citep{hirschberg75}

Levenshtein: \textit{metric} is a binary set. In algorithm True is matched with $0$, and False with $\infty$. Levenshtein algorithm accumulates / counts rather than finds. Levenshtein has to have addition to deal with lack of ordering of metric. This transformation depends on context of accumulation and position in word. Open question: the context of the accumulation solely originates from the position in the word. Is there metric for the words within a word to make this explicit and make it a search algorithm?


$f_\mathcal{C}: \mathcal{C} \times \mathcal{C} \mapsto \mathbb{N}$
\begin{equation}
    \label{eq:levenshtein_metric}
    \| a - b \|_\mathcal{C} = f_\mathcal{C}(a, b) = \begin{cases}
        0 & \text{if } a = b, \\
        1 & \text{else.}
    \end{cases}
\end{equation}

\begin{algorithm}
    \caption{Algorithm that maps a given distance matrix $d_{ij} = [p_i \neq q_j]$, where $[ \bullet ]$ is the Iverson bracket~\citep{knuth92}, to the Levenshtein distance of the character sequences $p \in \mathcal{C}^P$ and $q \in \mathcal{C}^Q$.}
    \begin{algorithmic}[1]
        \Function{levenshtein\_min}{$a: \mathbb{N}$, $x: \mathbb{N}$} $\to \mathbb{N}$
            \State \Return $\min\{ a + 1, x \}$
        \EndFunction
        \Statex
        \Function{levenshtein\_next}{$a: \mathbb{N}^Q$, $x: \mathbb{N}^{2 \times Q}$} $\to \mathbb{N}^Q$
            \State $a_{2 \ldots Q} \gets \min\{a_{1 \ldots Q-1} + x_{2,2 \ldots Q},\, a_{2 \ldots Q} + 1 \}$
            \State \Return \Call{levenshtein\_min$\backslash$}{$\min\{ x_{11} + x_{21},\, a_1 + 1 \},\, a_{2 \ldots Q}$}
        \EndFunction
        \Statex
        \Function{levenshtein\_distance}{$d: \mathbb{R}^{P \times Q}$} $\to \mathbb{N}$
            \State $\iota_{1 \ldots Q}: \mathbb{N}^{Q} \gets [1, \ldots, Q]$
            \State $\iota_{2 \ldots P}: \mathbb{N}^{P-1} \gets [2, \ldots, P]$
            \Statex
            \State $v_\mathrm{init}: \mathbb{N} \gets$ \Call{levenshtein\_min$\backslash$}{$\iota_{1 \ldots Q} + d_1$}
            \State \Return \Call{levenshtein\_next$/$}{$v_\mathrm{init},\, [\iota_{2 \ldots P} \;|\; d_{2 \ldots P}]$}
        \EndFunction
    \end{algorithmic}
\end{algorithm}

\section{Parallel Implementations}
\label{sec:parallel}
$\ldots$ schedule $\ldots$

\citep{avraham14}: The algorithms obtained for the discrete Fr\'echet distance with
shortcuts, run in time significantly better than those for the Fr\'echet
distance without shortcuts. 

\citep{agarwal14}: Computing the discrete Fr\'echet distance in subquadratic time. ($\overset{?}{\sim}$ linear search faster than more educated approaches on modern hardware if input data is small). The given \textit{subquadratic} complexity is only slightly better than quadratic.


Recently there has been some work on estimating Fr\'echet distances for uncertain curves. \citet{buchin23} shows that ... We note that in particular for the problem of finding approximations for the expected Fr\'echet distance when the vertices are either indecisive or even only implicitly given by their distribution function, our fast implementation might leverages MC approaches. + MC sampling perfectly befits the SIMD implementation! (For indecisive case: MC even allows assignment of probabilities to each point)
Work on Fr\'echet distance for uncertain curves, \textbf{e.g.}\footnote{there are more articles by Buchin about this, but there are not easy to cite} \citep{buchin23}: Given an uncertain curve and a second (certain or uncertain) curve, we seek to compute the lower and upper bound Fr\'echet distance, which are the minimum and maximum Fr\'echet distance for any realisations of the curves.
We prove that both problems are NP-hard for the Fr\'echet distance in several uncertainty models, and that the upper bound problem remains hard for the discrete Fr\'echet distance.
In contrast, the lower bound (discrete and continuous) Fr\'echet distance can be computed in polynomial time in some models.
Furthermore, we show that computing the expected (discrete and continuous) Fr\'echet distance is \#P-hard in some models.

